---
title: Predicting Rocket League Game Outcome Through Statistical Modeling
author:
  - name: Tanner Bessette
    affil: 
    orcid: 
  - name: |
      Faculty Advisor:&nbsp;
      Professor Higham
    affil: 
  - name: Department of Mathematics, Computer Science & Statistics
    
affiliation: 
  - num: 
    address: Department of Mathematics, Computer Science & Statistics
  - num: 
    address: St. Lawrence University
column_numbers: 3
logoright_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
logoleft_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png # this might be controlling posterdown logo in top right and left
output: 
  posterdown::posterdown_html:
    self_contained: true
bibliography: packages.bib
---













```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# What is Rocket League?

Rocket League is video game soccer played in cars. Each player controls their own car. Games last 5 minutes and the game-mode that I explored for the purposes of this project was the 3v3 mode.

# Project Goals

The main goal of my project was to train a model on professional Rocket League games and use that model to predict match outcomes (Win/Loss) for other professional Rocket League games. I wanted to get an idea of which variables in the dataset were most associated with winning games (besides obvious predictors like goals scored), and use these findings to figure out whether certain approaches/tactics that different professional teams utilize are more or less effective.

# Methods

I went into the project with an idea of two different statistical learning methods that I was going to attempt to predict game outcome with. The first was Lasso Regression, and the second was a Random Forest approach. After seeing the first few attempts at each of the models, it became clear that the Random Forest approach would be a much more effective modeling approach so that is what I focused the majority of my time on.

# What is Random Forest?

A random forest model is a type of statistical learning method that combines “decision trees” in order to make a prediction. Each tree is made up of a random subset of predictor variables from the full dataset. The predictions from each of these decision trees are combined to make one final prediction for each observation. 

# Results

In the end, I created a random forest model that predicted Rocket League game outcome with 86.35% accuracy. From every model that I created, there became a few clear predictor variables that were the strongest; teams with the most positioning time goalside of the ball and teams who got the most shots off during the game had a much better chance of winning those matches. I think that this helps reinforce the idea that constant pressure in the opponents' half remains the most effective strategy in Rocket League. 


```{r mytable, out.width='80%'}
knitr::kable(iris[1:10, 1:4], caption = 'Table caption.',align = 'c',"html")
```

Or with figures: Figure \@ref(fig:standard-plot), or Figure \@ref(fig:morefigs).

```{r standard-plot, out.width='80%', fig.align='center', fig.cap='Great figure!', fig.height=5}
plot(mtcars[1:2])
```

```{r morefigs, out.width='80%', echo=FALSE, fig.cap='Amazing, right?!', fig.height=5}
library(tidyverse)
library(tidymodels)
library(tree)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(knitr)
library(here)
library(vip)
library(caret)
RL_joined <- read_csv(here("RL_Datasets/RL_joined.csv"))
RL_numeric <- read_csv(here("RL_Datasets/RL_numeric.csv"), 
        col_types = cols(winner_factor = readr::col_factor()))

RL_numeric_tree_example <- RL_numeric |> 
                  select(c("winner_factor", 
                           "movement_time_high_air_diff",
                           "boost_time_zero_boost_diff",
                           "movement_time_slow_speed_diff",
                           "positioning_time_defensive_third_diff",
                           "positioning_time_offensive_half_diff"))
clean_tree <- rpart(winner_factor ~ ., 
                    data = RL_numeric_tree_example)
rpart.plot(clean_tree)
```

# Next Steps

Aliquam sed faucibus risus, quis efficitur erat. Vestibulum semper mauris quis tempus eleifend. Aliquam sagittis dictum ipsum, quis viverra ligula eleifend ut. Curabitur sagittis vitae arcu eget faucibus. In non elementum felis. Duis et aliquam nunc. Nunc pulvinar sapien nunc, vel pretium nisi efficitur in. Fusce fringilla maximus leo et maximus. Fusce at ligula laoreet, iaculis mi at, auctor odio. Praesent sed elementum justo. Aenean consectetur risus rhoncus tincidunt efficitur. Praesent dictum mauris at diam maximus maximus [@R-posterdown].

# Conclusion

Interestingly, although Rocket League coaches are constantly talking about keeping your car's movement speed high throughout the duration of the match, variables related to car speed had little to no effect on predicting game outcome.


```{r, include=FALSE}
knitr::write_bib(c('knitr','rmarkdown','posterdown','pagedown'), 'packages.bib')
```

# References
