---
title: "Start of Summer Research"
format: html
---
(With updated diff code from Prof. Higham)


Load in some necessary packages:
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(GGally)
library(here)
```

Load in the dataset:
```{r}
rl_df <- read.csv(here("RL_Datasets/games_by_teams.csv"))
```

View variable names in the dataset:
```{r}
names(rl_df)
```


(Next few plots before ggpairs are still from combined_df)
Ball Possession Plots:
```{r}
ggplot(data = combined_df, aes(x = winner, 
                         y = net_ball_poss)) +
  geom_boxplot()
```


Demo Plots:
```{r}
## This plot is very interesting, demo is not a statistic that I would've 
## expected to be so clear with the outliers. No team that dominated the demos 
## lost the game.
ggplot(data = combined_df, aes(x = winner, y = net_demo)) +
  geom_boxplot()
```

```{r}
## Really interesting that it seems like winning/losing the demo battle by
## 5 or less does not appear to have any sort of clear effect, but once it is 
## more than that it appears to have a very significant effect.
ggplot(data = combined_df, aes(x = winner, y = net_demo)) +
  geom_violin()
```

```{r}
ggplot(data = combined_df, aes(x = winner, y = net_demo)) +
  geom_jitter(width = 0.40)
```


Time behind ball plot:
```{r}
ggplot(data = combined_df, aes(x = winner, y = net_demo)) +
  geom_boxplot()
```

Want to see if the time behind ball correlates to time in defensive half:
```{r}
ggplot(data = combined_df, aes(x = winner, y = net_positioning_time_def_half)) +
  geom_boxplot()
```

It didn't, actually the more time spent in the defensive half appears to mean
the less likely it is to win the game. My best guess to what this means is that,
when you are constantly putting pressure on the other team, you are 
consistently behind the ball but you are also more likely at midfield or 
in the attacking half.


Net slow speed plots:
```{r}
ggplot(data = combined_df, aes(x = winner, y = net_slow_speed)) +
  geom_boxplot()
```

```{r}
ggplot(data = combined_df, aes(x = winner, y = net_slow_speed)) +
  geom_violin()
```

Supersonic speed splots:
```{r}
ggplot(data = combined_df, aes(x = winner, y = net_supersonic)) +
  geom_boxplot()
```

```{r}
ggplot(data = combined_df, aes(x = winner, y = net_supersonic)) +
  geom_jitter(width = 0.40)
```

I am very surprised to see that neither of the speed variables appear to be 
significant predictors. As a fan/player, keeping your speed high is one of the
number one things that typically gets attributed to winning/losing.


Shoot percentage plots:
```{r}
ggplot(data = combined_df, aes(x = winner, y = net_shoot_pct)) +
  geom_boxplot()
```

```{r}
ggplot(data = combined_df, aes(x = winner, y = net_shoot_pct)) +
  geom_violin()
```

```{r}
ggplot(data = combined_df, aes(x = winner, y = net_shoot_pct)) +
  geom_jitter(width = 0.40)
```
(jitter plots pretty much suck with this dataset except for looking at outliers)





PLOTS BELOW USING team_diff DATASET:
ggpairs plots to look at multiple predictors at a time:
net_ball_poss, net_shots, net_saves, net_assists
```{r}
ggpairs(data = team_diff, columns = c(4, 5, 7, 9, 10), 
        lower = list(combo = wrap(ggally_facethist, bins = 15)))
```
Net_ball_poss(6) appears to be a very weak predictor.
Net_saves(8) appears to be a fairly weak predictor of match outcome.
Net_shots(7) appear to be a moderate predictor of match outcome.
Net_assists(9) appear to be a strong predictor of match outcome, which is to be
expected.

net_ball_time_in_side, net_score, net_shoot_pct, net_bpm
```{r}
ggpairs(data = team_diff, columns = c(4, 6, 11, 12, 13), 
        lower = list(combo = wrap(ggally_facethist, bins = 15)))
```
Net_score(11) appears to have a strong correlation to match outcome.
Net_shoot_pct(12) also appears to have a fairly strong correlation to match 
outcome.
Net_ball_time_in_side(10) appears to have moderate correlation to match outcome.
Net_bpm(13) appears to be a weak predictor in match outcome.



net_boost_stolen, net_time_zero_boost, net_supersonic, net_slow_speed
```{r}
ggpairs(data = team_diff, columns = c(4, 17, 29, 36, 38), 
        lower = list(combo = wrap(ggally_facethist, bins = 15)))
```
Net_time_zero_boost(15) appears to have a slight correlation to match outcome.
Net_boost_stolen(14) appears to be a fairly weak predictor.
Net_supersonic(16) and net_slow_speed(17) appear to have almost no correlation 
to match outcome at all.
(ALL APPEAR TO BE WEAK)




net_ground_time, net_powerslide_time, net_positioning_time_def_half
```{r}
ggpairs(data = team_diff, columns = c("winner_binary", 
                                        "movement_time_powerslide_diff", 
                                        "positioning_time_defensive_half_diff",
                                        "movement_time_ground_diff"), 
        lower = list(combo = wrap(ggally_facethist, bins = 15)))
```
Net_ground_time(18) and net_powerslide(19) appear to be pretty weak predictors.
Net_positioning_time_def_half (20) appears to have correlation to match outcome.
(ALL WEAK, def_half best with corr of -0.162)


net_time_behind_ball, net_demo
```{r}
ggpairs(data = team_diff, columns = c("winner_binary", 
                                        "demo_inflicted_diff",
                                        "positioning_time_behind_ball_diff"), 
        lower = list(combo = wrap(ggally_facethist, bins = 15)))
```
Both net_time_behind_ball(21) and net_demo(22) appear to be good predictors.
Demo_inflicted_diff has weak "corr" but obvious visual diff.
(tim_behind_ball_diff has corr: 0.666)


Variables with corr values above 0.5: core_score_diff, 
core_shooting_percentage_diff, core_assists_diff, positioning_time_behind_ball
CLOSE/potentially strong predictors: core_shots_diff, demo_inflicted_diff,
ball_time_in_side_diff (corr 0.25)

Further investigating the "potentially strong" ones:
core_shots_diff:
```{r}
ggplot(data = team_diff, aes(x = winner, 
                             y = core_shots_diff)) +
  geom_boxplot()
```
In my opinion, there is a clear correlation.


demo_inflicted_diff:
```{r}
ggplot(data = team_diff, aes(x = winner, y = demo_inflicted_diff)) +
  geom_boxplot()
```
If demo difference is big, there appears to be a very clear correlation to 
game outcome. However, when the demo difference is close, it appears to be a 
very weak predictor.


ball_time_in_side_diff:
```{r}
ggplot(data = team_diff, aes(x = winner, y = ball_time_in_side_diff)) +
  geom_boxplot()
```
There appears to be a slight correlation between ball_time_in_side_diff and 
match outcome.


positioning_time_behind_ball_diff plot:
```{r}
ggplot(data = team_diff, aes(x = winner, 
                             y = positioning_time_behind_ball_diff)) + 
  geom_boxplot(fill="coral1",colour ="black")
```


core_shooting_percentage_diff
```{r}
ggplot(data = team_diff, aes(x = winner, 
                             y = core_shooting_percentage_diff)) + 
  geom_boxplot(fill="coral1",colour ="black")
```


Plot to use in presentation on Thurs 6/22:
```{r}
ggpairs(data = team_diff, columns = c("winner", 
                                      "demo_inflicted_diff",
                                      "positioning_time_behind_ball_diff",
                                      "core_shots_diff",
                                      "boost_time_zero_boost_diff"), 
        lower = list(combo = wrap(ggally_facethist, bins = 15)))
```




